
---
title: "Problem Set 02 Solutions"
output:
  html_document:
    theme: lumen
---

```{r message = F}
library(dplyr)
library(ggplot2)
library(readr)
```


```{r, message = F}
nc <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vTm2WZwNBoQdZhMgot7urbtu8eG7tzAq-60ZJsQ_nupykCAcW0OXebVpHksPWyR4x8xJTVQ8KAulAFS/pub?gid=202410847&single=true&output=csv")
```


```{r eval = F}
glimpse(nc)
```

# Exercise 1

What type of variable is R considering the variable `habit` to be? What variable type is `visits`? (answer with text)

**Answer: habit is listed as <chr> (character), and visits is an integer <int>**


```{r eval = F}
ggplot(data = nc, aes(x = weeks, y = weight)) + 
  geom_point() + 
  labs(x = "Length of pregnancy (in weeks)", y = "Birth weight of baby (lbs)", 
       title = "Relationship between pregnancy duration and newborn weight")
```

## Exercise 2
1. Is there a positive or negative relationship between these variables? (text only to answer)

**Answer: positive**

## Exercise 3
Make a graph showing `weeks` again on the x axis and the variabel `gained` on the y axis (the amount of weight a mother gained during pregnancy).  Include axis labels with measurement units, and a title. (code only to answer)

```{r fig.height=3, fig.width=3}
ggplot(data = nc, aes(x = weeks, y = gained))+ 
  geom_point()
```

## Exercise 4
Study the code below, and the resulting graphical output. 

**A.** What did adding the argument `color = premie` accomplish? 

**Answer: the color of the points corresponds to whether a birth was a premie or a full term birth**
     
**B.** How many **variables** are now displayed on this plot?  

**Answer: 3 variables are shown**
     
**C.** What appears to (roughly) be the pregnancy length cutoff for classifying a newborn as a "premie"" versus a "full term". 

**Answer: anywhere between 36 and 38 seems reasonable**

```{r fig.height=3, fig.width=5, warning=F}
ggplot(data = nc, aes(x = weeks, y = gained, color = premie))+ 
  geom_point() + 
  labs(x = "Pregnancy length (wks)", y = "Maternal weight gain (lbs)")
```

## Exercise 5
Make a new scatterplot that shows a mothers age on the x axis (variable called `mage`) and birth weight of newborns on the y axis (`weight`). Color the points on the plot based on the gender of the resulting baby (variable called `gender`). Does there appear to be any strong relationship between a mother's age and the weight of her newborn? (code and text to answer)

**Answer: there does not appear to be any strong relationship between the mother's age and the weight of her newborn...the points are in a big blob** 

```{r fig.height=3, fig.width=5}
ggplot(data = nc, aes(x = mage, y = weight, color = gender)) + 
  geom_point() + 
  labs(x = "Mother's age (in years)", y = "Mother's weight gain during pregnancy (lbs)", 
       title = "Weight gain of mothers by age and baby gender")
```


```{r fig.height=3, fig.width=5, warning = F}
ggplot(data = nc, aes(x = weeks))+ 
  geom_histogram(binwidth = 1, color = "white", fill = "steelblue")
```

## Exercise 6
Inspect the histogram of the `weeks` variable. Answer each of the following with **text**. 

**A.** The y axis is labeled **count**. What is specifically being counted in this case? Hint: think about what each case is in this data set. 

**Answer: the number of births that fall into each bin specified on the histogram**
    
**B.** What appears to be roughly the average lenght of pregnancies in weeks?

**Answer: 39 weeks...37, 38, or 40 also acceptable**
    
**C.** If we changed the binwidth to 100, how many bins would there be? Roughly how many cases would be in each bin?
    
**Answer: there would be one bin. It would contain every case...so 800 births would be in it**

## Exercise 7
Make a histogram of the birth `weight` of newborns (which is in lbs), including a title and axis labels. (code only to answer)

```{r fig.height=3, fig.width=5, warning = F}
ggplot(data = nc, aes(x =  weight)) +
  geom_histogram(binwidth = 1, color = "white", fill = "sienna") + 
  labs(x = "Newborn weight (lbs)", y = "Count", title = "Distribution of newborn birth weights")
  
```

## Exercise 8
Make a histogram of newborn birth `weight` split by `gender` of the child. Set the binwidth to 0.5. Which gender appears to have a slightly larger average birth weight? (code and text to answer)

**Answer: Males have a slightly higher average birth weight**

```{r fig.height=3, fig.width=5, warning = F}
ggplot(data = nc, aes(x = weight)) +
  geom_histogram(binwidth = 0.5, color = "white", fill = "steelblue") +
  facet_wrap(~ gender, ncol = 1)
```

## Exercise 9
Make a boxplot of the weight `gained` by moms, split by the maturity status of the mothers (`mature`).  Include axis labels and a title on your plot. Is the **median** weight gain during pregnancy larger for younger or older moms? (text and code)

**Answer it is slighly greater for younger moms**

```{r fig.height=5, fig.width=5, warning = F}
ggplot(data = nc, aes(x = mature, y = gained)) + 
  geom_boxplot(fill = "steelblue") + 
  labs(x = "Maturity status of mother", y = "Pregnancy weight gain (lbs)", 
       title = "Pregnancy weight gain for older and younger mothers")
```

## Exercise 10
Make a boxplot of pregnancy duration by smoking `habit`. Is the duration of pregnancy more **variable** for smokers or non-smokers? (i.e. which group has the greater spread for the variable `weeks`?). (code and text to answer)

**answer: based on the RANGE, pregnancy duration is more variable for nonsmokers...based on the IQR pregnancy duration is more variable for smokers**

```{r fig.height=3, fig.width=5, warning = F}
ggplot(data = nc, aes(x = habit, y = weeks)) + 
  geom_boxplot(fill = "steelblue")
```


# Independent Practice 
For the following, you need to determine which type of plot to use, **make the plot**, and answer any questions with **text**. There is a table at the end of this document that can help you determine which plot to use, given the question/types of variables. 

## Exercise 11
Using a data visualization, visually assess: Is the variable for father's age (`fage`) symmetrical, or does it have a skew?
```{r fig.height=3, fig.width=5, warning = F}
ggplot(data = nc, aes(x = fage)) + 
  geom_histogram(binwidth = 2)
```

**Answer: there is a right skew. Note to the grader...technically a boxplot could also be used to assess this...though we did not show the students how** 

## Exercise 12
Using a data visualization, visually assess: (in this sample) is the median birth `weight` of babies greater for white or non-white mothers (variable called `whitemom`)?

**Answer: median weight of babies  is greater for white mothers** 

```{r fig.height=3, fig.width=5, warning = F}
ggplot(data = nc, aes(x = whitemom, y = weight)) +
  geom_boxplot(fill = "lightblue") 
```

## Exercise 13 
Using a data visualization, visually assess: (in this sample) as a mother's age increases, does the duration of pregnancy appear to decrease? 

*Answer: no...there does not appear to be much of a relationship** 
```{r fig.height=3, fig.width=5, warning = F}
ggplot(data = nc, aes(x = mage, y = weeks)) +
  geom_point()
```
