
---
title: "SDS/MTH 220 Project Data Proposal"
author: "WRITE YOUR GROUP NAME HERE"
date: "Last updated on `r Sys.Date()`"
output:
  html_document:
    fig_caption: yes
    theme: lumen
    toc: yes
    toc_depth: 2
    df_print: kable
    toc_float:
      collapsed: no
---

```{r, message=FALSE}
# This sets the options for the output of each code chunk in the resulting 
# .html report:
knitr::opts_chunk$set(
  echo = TRUE, # Do not show code in each code chunk
  message = FALSE, # Hide all messages
  warning = FALSE, # Hide all warnings
  fig.width = 16/2, # Set the widths of all figures/visualizations
  fig.height = 9/2 # Set the height of all figures/visualizations
)

# Load all packages here:
library(readr)
library(dplyr)
library(ggplot2)
library(janitor)

# Set seed value of random number generator to get "replicable" random numbers.
# The choice of seed value of 76 was an arbitrary one on my part.
set.seed(76)
```



# Big-picture

## Research question

_What is your research question?_



## Description of data

_Please give a very short description of the data set along with it's original source._


## Load data into R

_Include the code to load your data here. If your data is not confidential nor is it private in nature, consider publishing it as a `.csv` file on Google Sheets as in the code chunk below; instructions on how to do this are in Steps 1-6 [here](https://twitter.com/rudeboybert/status/1055821846539591680). If the data shouldn't be published online, then please submit the spreadsheet file on Moodle._

```{r}

```


## Clean variable names

_The `clean_names()` function from the `janitor` package will clean your variable names, making them easier to work with._

```{r}

```



***



# Variables

## Identification variable

_What is your identification (ID) variable (if you have one)?_


## Outcome variable

_What is your outcome variable $y$? What are its units of measurement?_



## Numerical explantory variable

_What is your numeric explanatory variable? What are its units of measurement?_



## Categorical explantory variable

_What is your categorical explanatory variable? Does it have between 3 and 5 levels. Please list the different levels._

```{r}

```



***


# Rows/observations

## Observational units

_What is the observational unit of your data? In other words, what does each row in your data represent?_


## Sample size

_How many rows/cases are in the data i.e. what is the sample size? Is the sample size at least 50?_


```{r}

```




***


# Preview of data

## Pare down variables

_`select()` the following variables and drop all others:_

1. The identification variable
1. The outcome variable $y$
1. The numerical explanatory variable
1. The categorical explanatory variable
1. Optional: any other variable you find interesting

```{r}

```


## Preview data

_Display a random sample of 5 rows of your data frame. You'll get the same 5
rows everytime you knit this document and hence replicable results because we
set the seed value of the random number generator in the first code chunk above._

```{r}

```





