
---
title: "SDS/MTH 220 Data Proposal Example"
author: "Alexis Cohen, Andrianne Dao, Isabel Gomez, Jenny Smetzer, and Albert Y. Kim"
date: "Last updated on `r Sys.Date()`"
output:
  html_document:
    fig_caption: yes
    theme: lumen
    toc: yes
    toc_depth: 2
    df_print: kable
    toc_float:
      collapsed: no
---

```{r, message=FALSE}
# Load all packages here:
library(readr)
library(dplyr)
library(ggplot2)
library(janitor)

# Set seed value of random number generator to get "replicable" random numbers.
# The choice of seed value of 76 was an arbitrary one on my part.
set.seed(76)
```

# Please note

**Your data proposal may not follow the exact same steps in the example below. Jenny and Albert are merely providing one example to give you an overall sense of where you are heading; the details may change from project to project.**



***



# Big-picture

## Research question

_What is your research question?_

Overall, we are hoping to better understand how and if school conditions influence student performance. Our question specifically is whether percentage of economically disadvantaged students and school size predict the average math SAT score in Massachusetts high schools. 

## Description of data

_Please give a very short description of the data set along with it's original source._

We have data from all schools in Massachusetts from 2017. The data set contains SAT scores, and some information about school demographics. While the data was downloaded from Kaggle [here](https://www.kaggle.com/ndalziel/massachusetts-public-schools-data), the original source of the data are Massachusetts Department of Education [reports](http://profiles.doe.mass.edu/state_report/).

## Load data into R

_Include the code to load your data here. If your data is not confidential nor is it private in nature, consider publishing it as a `.csv` file on Google Sheets as in the code chunk below; instructions on how to do this are in Steps 1-6 [here](https://twitter.com/rudeboybert/status/1055821846539591680). If the data shouldn't be published online, then please submit the spreadsheet file on Moodle._

```{r, message=FALSE}
ma_schools <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSrWSNyNqRVA950sdYa1QazAT-l0T7dl6pE5Ewvt7LkSm9LXmeVNbCbqEcrbygFmFyK4B6VQQGebuk9/pub?gid=1469057204&single=true&output=csv")
```


## Clean variable names

_Piping your data frame into the `clean_names()` function from the `janitor` package will clean your variable names, making them easier to work with._

```{r}
ma_schools <- ma_schools %>% 
  clean_names()
```


## Explore your data

_Be sure to explore your data. Note that `eval=FALSE` is set so that R Markdown doesn't "evaluate" this code chunk, i.e. it will ignore it in the ultimate `.html` report. You should run this code on your own, but not in the ultimate `.html` report._

```{r, eval = FALSE}
glimpse(ma_schools)
```



***



# Variables

## Identification variable

_What is your identification (ID) variable (if you have one)?_

Our identification variable is `school_name`; it allows us to uniquely identify each school.

## Outcome variable

_What is your outcome variable $y$? What are its units of measurement?_

The outcome variable is called `average_sat_math`: the average Math SAT score for a school measured in points.  We did not have data on overall SAT scores for each school, so we decided to focus solely on the math SAT score in order to make this report more concise.

## Numerical explantory variable

_What is your numeric explanatory variable? What are its units of measurement?_

The numerical explantory variable is called `percent_economically_disadvantaged` and its unit of measurement are percentage points between 0-100.  It represents the percentage of students in a school that are considered "economically disadvantaged".

## Categorical explantory variable

_What is your categorical explanatory variable? Does it have between 3 and 5 levels. Please list the different levels._

Since we didn't feel any of the existing categorical variables in `ma_schools` were appropriate for our analysis, we hope to create a new categorical explantory variable called `size`, which will be based on the numerical variable `total_enrollment`. We'll divide this numerical variable into three equally sized groups, and hence our levels will be: `small`, `medium`, and `large`. We will do this in the next step of the project once we are more experienced with data wrangling.

<!--
We constructed this variable by cutting the numeric variable `total_enrollment` into `n=3` roughly equally sized parts so that schools with 0-341 students were called `small`, schools with 342-541 students were classified as `medium` and those with > 541 students were `large`. 
-->

```{r, eval=FALSE, echo=FALSE}
ma_schools <- ma_schools %>% 
  mutate(
    school_size = cut_number(total_enrollment, n = 3),
    size = recode_factor(
      school_size, "[0,341]" = "Small", "(341,541]" = "Medium", "(541,4.26e+03]" = "Large"
    )
  )
```



***


# Rows/observations

## Observational units

_What is the observational unit of your data? In other words, what does each row in your data represent?_

Each row/case in our data set will be a high school in Massachusetts (see below):

## Sample size

_How many rows/cases are in the data i.e. what is the sample size? Is the sample size at least 50?_

For now there are 1861 rows in our original data set `ma_schools`. However, eventually we would like to limit our analysis to only high schools as younger grades do not take the SAT. We will do this in the next step of the project once we are more experienced with data wrangling.

<!--
```{r, eval=FALSE, echo=FALSE}
ma_schools <- ma_schools %>% 
  filter(x11_enrollment > 0 & x12_enrollment > 0)
```

There also appear to be 58 `NA` missing values for `average_sat_math`:

```{r, eval=FALSE, echo=FALSE}
ma_schools %>% 
  mutate(is_missing = is.na(average_sat_math)) %>% 
  summarize(total_missing = sum(is_missing))
```

We will drop these rows while being sure to report this fact in the next step of the project, as this may have implications on our interpretation of the results. So the final sample size will be 332. 


```{r, eval=FALSE, echo=FALSE}
ma_schools <- ma_schools %>% 
  filter(!is.na(average_sat_math))
```
-->


***


# Preview of data

## Pare down variables

_`select()` the following variables **in this order** and drop all others. Eliminating all unnecessary variables will making visually exploring the raw values less taxing mentally, as we'll have less data to look at._

1. _The identification variable_
1. _The outcome variable $y$_
1. _The numerical explanatory variable_
1. _The categorical explanatory variable_
1. _Optional: any other variable you find interesting_

Recall that we didn't find a categorical variable to our liking in the original `ma_schools` data. So eventually we'll create a categorical explanatory variable `school_size` with three levels: `small`, `medium`, `large` based on the values of `total_enrollment`. We will do this in the next step of the project once we are more experienced with data wrangling.

```{r}
ma_schools <- ma_schools %>% 
  select(school_name, average_sat_math, percent_economically_disadvantaged, total_enrollment)
```


## Preview data

_Display a random sample of 5 rows of your data frame by piping it into the `sample_n(5)` function from the `dplyr` package . You'll get the same 5 rows everytime you knit this document and hence replicable results because we set the seed value of the random number generator in the first code chunk above._

We see we have missing values below. Some of these are missing because some of the schools are not high schools so the students don't take the Math SAT. We will address these missing values in the next step of the project once we are more experienced with data wrangling.

```{r}
ma_schools %>% 
  sample_n(5)
```





